Eurovision Song Contest: semifinal 1, 2009

### 2019-12-23 inst/eurovision.Rmd
work

---
title: "Eurovision Song Contest"
output: html_document
---

@@

Each row corresponds to a competitor and each column to a judge.  Note
that the matrix is not square: the last two columns correspond to
Germany and the UK, who were judges but not competitors.  I have
removed the first column in the Wikipedia table (which gave the total
of the points), replaced self-voting entries with `NA`, and replaced
blanks with `00`; every entry is padded to two digits.

@@

Now, take the first row of 'preference'.  This represents the votes
cast *by* (sic) Montenegro (`"ME"`).  Their favourite was [last column]
Bosnia & Herzegovina (`"BH"`), who they gave rank 1 to.  Their second
favourite was Macedonia, their third was Turkey, and so on.  They were
not allowed to vote for themselves, which is why the first column is
`NA`.  So the order was: `bh`, `mc`, `tu`, `ic`, `ro`, `is`, `ar`,
`fi`, `br`, `ma`.

@@

```{r,plotdiffstrength}
plot(rowSums(wiki_matrix,na.rm=TRUE),maxp(euro2009),
     pch=16,xlab="points scored",ylab="MLE strength")
```

>>>>

---
title: "Eurovision Song Contest: semifinal 1, 2009"
output: html_document
---

@@

Each row corresponds to a competitor and each column to a judge.  Note
that the matrix is not square: the last two columns correspond to
Germany and the UK, who were judges but not competitors.  I have
removed the first column in the Wikipedia table (which gave the total
of the points), replaced self-voting entries with `NA`, and replaced
blanks with `00`; every entry is padded to two digits.  Thus the first
row corresponds to votes cast for `ME`, Montenegro.  We see that `BE`
(Belgium) awarded them three points, `am` (Armenia) five points, and
so on.

@@

Now, take the first row of 'preference'.  This represents the votes
cast *by* (sic) Montenegro (`"ME"`).  Their favourite was [last
column] Bosnia & Herzegovina (`"BA"`), who they gave rank 1 to.  Their
second favourite was Macedonia, their third was Turkey, and so on.
They were not allowed to vote for themselves, which is why the first
column is `NA`.  So the order was: `BA` (1), `MK` (2) , `TR` (3), `IS`
(4), `RO` (5), `IL` (6), `AM` (7), `FI` (8), `BY` (9), `MT` (10).  The
other countries (`CZ`, `BE`, `SW`, `CH`, `BG`, `PT`) did not receive
any points from Montenegro.

We need to convert matrix `preference` into a likelihood function,
here `euro2009`:

@@

```{r,orderplot}
m <- maxp(euro2009)  # likelihoods
points_scored <- rowSums(wiki_matrix,na.rm=TRUE)  # points
ox <- rank(-points_scored)
oy <- rank(-m)
par(pty='s') # square plot
plot(ox,oy,asp=1,pty='s',xlim=c(0,18),ylim=c(0,18),pch=16,
xlab="official order",ylab="my order",main="Eurovision 2009")
par(xpd=TRUE) # allow country names to appear outside plotting region
for(i in seq_along(ox)){text(ox[i],oy[i],names(m)[i],pos=4,col='gray') }
par(xpd=FALSE) # stop diagonal line from protruding beyond plotting region
abline(0,1)
```

In the graph above, we see that both systems agree on the first four places.
We can also plot the numerical values:

```{r}
plot(log(rowSums(wiki_matrix,na.rm=TRUE)),log(maxp(euro2009)),
     pch=16,xlab="points scored",ylab="MLE strength")
```
